### 2017-12-28 article.tex
bibliography

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Topic classification in Portuguese news articles\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in
Xplore and should not be used} \thanks{Identify applicable funding
agency here. If none, delete this.} }

\author{\IEEEauthorblockN{André F. Santos}
\IEEEauthorblockA{\textit{CRACS \& INESC-Porto LA} \\ \textit{Faculty
of Sciences, University of Porto}\\ Porto, Portugal \\
afs@inesctec.pt} }

\maketitle

\begin{abstract} This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper
[title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special
Characters, Footnotes, or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords} topic classification, machine learning, text
mining \end{IEEEkeywords}

\section{Introduction}
The last decades have witnessed a fast increase on the rate of publication
of digital text documents. Traditional document types, such as news
articles, scientific papers or books are now published online together
with new formats, such as blog posts or tweets, each having thousands
or millions of new documents published each day.

And it was not only the publication step which has moved to the
digital world; in fact, most often nowadays the whole document
pipeline happens digitally, with virtual tools available for
preparing, writing, styling, publishing and sharing.

Having the entire workflow happening within the digital world presents
some opportunities when compared to the traditional process. In
particular, due to todays processing power, tasks related to the
manipulation of the information contained within these documents
(searching, compiling, annotating, sharing, \dots) can now be
performed automatically and to a large amount of articles.

In addition to the document content (for example, in a news article,
the \emph{title}, \emph{lead} and \emph{body}), its metadata is also
important: author(s), date of publication, source, topic, mentioned
entities and their relations, etc. Some of this metadata might be
filled in and stored along with the document (e.g. \emph{author} and
\emph{date of publication}); other is usually extracted from the
document content (e.g. mentioned entities).

Online news articles first appeared as reprints from traditional
newspapers; nowadays, however, they represent now the primary source
of news for some segments of the population, both in developed and
developing countries (whether consumed directly in the newspaper
website, or indirectly (e.g. through a social media application or a
feed catcher).

Unofficially know as \emph{the fourth branch of government}, the press
plays a vital role within our society, keeping us informed regarding
the current state of affairs (at a local and global scale) and acting
as a watchdog for the other three powers (legislative, executive and
judicial). The (lack of) freedom of press and access to the news in a
given country is even often considered an indicator of a lack of
democracy\cite{goode2009social}.

As such, improving the ways citizens can access the information (view
it, query it and search it) contained in news articles has the
potential to contribute for a more informed society and, ultimately, a
better society.

An example of a feature which improves information access is the
categorization of news articles by the topic (or topics) of its
content. Having such a categorization may influence the way the
information is stored, organized, displayed and queried.

The simplest way of achieving this categorization is to have the
author of the article to manually introduce it (e.g. the journalist
typing it on the news article authoring framework); however, this
solution presents some challenges:
\begin{itemize}
    \item it increases the amount of work the author has to do
    \item the author might not be sure which categories are available
    \item the author might not be sure which category is the best
        (e.g. \emph{Economy} vs  \emph{Finance}
    \item it might be impractical if the goal is to categorize an
        existing (large) corpus
\end{itemize}

Thus, an automated way of categorizing news articles could solve some
of these problems and decrease the burden of this task.

The challenges of document classification have been well studied
within the machine learning research field of study. Given a corpus of
already classified documents, several algorithms might be applied to
train a classifier capable of determining the category of additional
articles.

\section{Methods}
% 500 articles from each category to train
% 200 articles to evaluate

\subsection{Classification algorithms}
\subsubsection{Decision tree}
\subsubsection{k-nearest neighbors}
\subsubsection{Naive Bayes}
\subsubsection{Neural network}
\subsubsection{Support vector machine}

\subsection{Most informative terms}
% for real-time topic suggestion (for an IDE for example)
% memory limit
% 100 terms? 1000 terms?

\section{Results}
\subsection{Evaluation}
\subsection{Limited lexicon}

\section{Discussion}

\subsection{Maintaining the Integrity of the Specifications}

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
%\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}
\bibliography{article}
\bibliographystyle{splncs03}

\end{document}

### 2018-01-02 article.tex
Dataset description in Methods

@@
 \usepackage{textcomp}
+\usepackage{listings}
+\usepackage{hyperref}
+\usepackage[T1]{fontenc}
+\usepackage[utf8x]{inputenc}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
-\title{Topic classification in Portuguese news articles\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in
-Xplore and should not be used} \thanks{Identify applicable funding
-agency here. If none, delete this.} }
+\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true,literate={á}{{\'a}}1 {ê}{{\^e}}1 {é}{{\'e}}1 }
+
+\title{Topic categorization in Portuguese news articles\\
+\thanks{Fundação para a Ciência e Tecnologia} }
 
@@
 
-\begin{IEEEkeywords} topic classification, machine learning, text
+\begin{IEEEkeywords} topic categorization, machine learning, text
 mining \end{IEEEkeywords}
@@
 digital world; in fact, most often nowadays the whole document
-pipeline happens digitally, with virtual tools available for
+lifecycle happens digitally, with virtual tools available for
 preparing, writing, styling, publishing and sharing.
@@
 (searching, compiling, annotating, sharing, \dots) can now be
-performed automatically and to a large amount of articles.
+performed automatically and targetting a large amount of articles.
 
@@
 categorization of news articles by the topic (or topics) of its
-content. Having such a categorization may influence the way the
-information is stored, organized, displayed and queried.
+content. The presence of such a categorization may influence the way
+the information is stored, organized, displayed and queried.
 
@@
     \item the author might not be sure which category is the best
-        (e.g. \emph{Economy} vs  \emph{Finance}
-    \item it might be impractical if the goal is to categorize an
+        (e.g. \emph{Economics} vs  \emph{Finance})
+    \item it dows not scale -- e.g. if the goal is to categorize an
         existing (large) corpus
@@
 \section{Methods}
-% 500 articles from each category to train
-% 200 articles to evaluate
+\subsection{The dataset}
+We gathered a dataset of news articles published in
+\textit{Oservador}\footnote{\url{http://observador.pt}}, one of the
+main Portuguese newspapers, and which stands out from the others for
+being fairly young (it was created in May 2014) and for existing
+exclusively online.
+
+We gathered all the categories used by Observador, and ordered them
+from the most common to the least common. We selected the ones which
+had more than 1.000 articles in our dataset, and reduced our original
+dataset to include only articles from these categories, which
+were\footnote{Category names translated to English by the authors.}:
+\begin{enumerate}
+    \item International
+    \item National
+    \item Economics
+    \item Politics
+    \item Sports
+    \item Science
+    \item Culture
+    \item Music
+\end{enumerate}
+
+We them randomly selected, from each category, 700 articles to be used
+to train the classifiers, and 200 to be used to evaluate their
+performance.
+
+For each article, we had available its contents (title, lead, body)
+and several metadata fields (publication date, category, tags, etc). A
+truncated JSON representation of an article can be found in
+Listing~\ref{lst:article}.
+
+\begin{lstlisting}[caption={Example of JSON representation of an article},label={lst:article}, extendedchars=true]
+{
+  "Type" : "sapo.obj.creativework.article",
+  "Source" : {
+      "Name" : "Observador"
+  },
+  "Pretitle" : "Benfica",
+  "Title" : "Ruben Amorim com rotura total do ligamento cruzado",
+  "Author" : {
+      "Name" : "Observador"
+  },
+  "Tags" : [
+      "benfica",
+      "desporto",
+      "futebol",
+      "ruben amorim"
+  ],
+  "PublishDate" : ISODate("2014-08-25T18:33:00Z"),
+  "Lead" : "Depois de Fejsa, mais uma baixa. O internacional português [...]",
+  "Body" : "<p>O pior cenário confirmou-se. O Benfica informou esta segunda-feira [...]",
+  "URL" : "http://observador.pt/2014/08/25/ruben-amorim-com-rotura-total-ligamento-cruzado/",
+  "CategoryPaths" : [
+      "Desporto"
+  ],
+  "Domain" : "observador.pt",
+  "Language" : "pt_PT",
+  ...
+}
+\end{lstlisting}
+
+\subsection{Preprocessing the articles}
 
@@
 \section{Results}
+
+\begin{table}[htbp]
+    \caption{Categories confusion matrix (k-nearest neighbors)}
+\begin{center}
+\begin{tabular}{l|rrrrrrrr}
+              & ciencia     & cultura     & desporto     & economia     & mundo       & musica       & pais        & politica \\\hline
+ciencia       & \textbf{64} & 18          & 25           & 29           & 21          & 37           & 4           & 2 \\
+cultura       & 4           & \textbf{37} & 16           & 9            & 19          & 107          & 3           & 5 \\
+desporto      & 1           & 4           & \textbf{130} & 8            & 19          & 35           & 2           & 1 \\
+economia      & 0           & 6           & 9            & \textbf{127} & 12          & 32           & 1           & 13 \\
+mundo         & 9           & 8           & 25           & 31           & \textbf{82} & 32           & 1           & 12 \\
+musica        & 0           & 6           & 1            & 1            & 0           & \textbf{180} & 0           & 1 \\
+pais          & 8           & 15          & 24           & 42           & 17          & 64           & \textbf{14} & 16 \\
+politica      & 1           & 8           & 7            & 36           & 19          & 24           & 3           & \textbf{102} \\
+
+\end{tabular}
+\label{tab1}
+\end{center}
+\end{table}
+% >\\  error.rate.knn
+% [1] 0.5368156
 \subsection{Evaluation}
@@
 \section{Discussion}
-
-\subsection{Maintaining the Integrity of the Specifications}
-
-The IEEEtran class file is used to format your paper and style the text. All margins, 
-column widths, line spaces, and text fonts are prescribed; please do not 
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement 
-and others are deliberate, using specifications that anticipate your paper 
-as one part of the entire proceedings, and not as an independent document. 
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a 
-separate text file. Complete all content and organizational editing before 
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been 
-formatted and styled. Do not number text heads---{\LaTeX} will do that 
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text, 
-even after they have been defined in the abstract. Abbreviations such as 
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-%\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
+% nested categories
+% or not mutex categories
 
@@
 
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
+André Santos has a scholarship from Fundação para a
+Computação Científica Nacional.
 
-\section*{References}
 \bibliography{article}

### 2018-01-03 article.tex
Preprocessing

@@
         (e.g. \emph{Economics} vs  \emph{Finance})
-    \item it dows not scale -- e.g. if the goal is to categorize an
+    \item it does not scale -- e.g. if the goal is to categorize an
         existing (large) corpus
@@
 
+In this article, we describe the preliminary results obtained in
+developing a classifier to categorize news articles, and in developing
+a similar classifier with restrictions added to the maximum size of
+the lexicon used % TODO
+
 \section{Methods}
+%TODO
+we first needed to choose and obtain a suitable dataset, and then
+cleaning and preparing it to be used to train the classifiers.
+
 \subsection{The dataset}
 We gathered a dataset of news articles published in
-\textit{Oservador}\footnote{\url{http://observador.pt}}, one of the
+\textit{Observador}\footnote{\url{http://observador.pt}}, one of the
 main Portuguese newspapers, and which stands out from the others for
 being fairly young (it was created in May 2014) and for existing
-exclusively online.
+exclusively online. The initial dataset comprised 42.475 entries, from
+which we used only a subset, for reasons later described.
 
@@
 had more than 1.000 articles in our dataset, and reduced our original
-dataset to include only articles from these categories, which
-were\footnote{Category names translated to English by the authors.}:
-\begin{enumerate}
-    \item International
-    \item National
-    \item Economics
-    \item Politics
-    \item Sports
-    \item Science
-    \item Culture
-    \item Music
-\end{enumerate}
+dataset to include only articles from these categories.
+Figure~\ref{fig:articles} presents an overview of the selected
+categories and the number of articles available for each one.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.97\columnwidth]{imgs/articles}}
+\caption{Total number of articles retrieved for each category}
+\label{fig:articles}
+\end{figure}
 
@@
 {
-  "Type" : "sapo.obj.creativework.article",
-  "Source" : {
-      "Name" : "Observador"
+  Type: "sapo.obj.creativework.article",
+  Source: {
+      Name : "Observador"
   },
-  "Pretitle" : "Benfica",
-  "Title" : "Ruben Amorim com rotura total do ligamento cruzado",
-  "Author" : {
-      "Name" : "Observador"
+  Pretitle: "Benfica",
+  Title: "Ruben Amorim com rotura total do ligamento cruzado",
+  Author: {
+      Name: "Observador"
   },
-  "Tags" : [
+  Tags: [
       "benfica",
@@
   ],
-  "PublishDate" : ISODate("2014-08-25T18:33:00Z"),
-  "Lead" : "Depois de Fejsa, mais uma baixa. O internacional português [...]",
-  "Body" : "<p>O pior cenário confirmou-se. O Benfica informou esta segunda-feira [...]",
-  "URL" : "http://observador.pt/2014/08/25/ruben-amorim-com-rotura-total-ligamento-cruzado/",
-  "CategoryPaths" : [
+  PublishDate: ISODate("2014-08-25T18:33:00Z"),
+  Lead: "Depois de Fejsa, mais uma baixa. O internacional português [...]",
+  Body: "<p>O pior cenário confirmou-se. O Benfica informou esta segunda-feira [...]",
+  URL: "http://observador.pt/2014/08/25/ruben-amorim-com-rotura-total-ligamento-cruzado/",
+  CategoryPaths: [
       "Desporto"
   ],
-  "Domain" : "observador.pt",
-  "Language" : "pt_PT",
+  Domain: "observador.pt",
+  Language: "pt_PT",
   ...
@@
 
+Originally, the dataset was obtained as large (more than 2.5 million
+entries) MongoDB collection, containing articles from several
+Portuguese and international newspapers. The process required to
+transform this collection into data our classifiers could process
+required: querying the database, exporting the news articles and
+splitting them into a train and an evaluation datasets.
+
+The database query selected articles from \texttt{Observador} where
+the body had a length greater than 100 characters (to
+discard some malformed articles which had an empty body or a body
+composed of only a few words), and the categories included at least
+one of the categories we previously selected.
+
+For each article returned by the query, the pretitle, title, subtitle
+and lead fields, if present, were simply copied to a plain text file,
+separated by blank lines. The body field, however, was stored in the
+database in HTML format. As such, the HTML tags had to be stripped,
+and then it was also added to the plain text file.
+
+The files were then stored in folders, separated by the category.
+Furthermore, for each category, 700 articles were allocated to the
+train set, and 200 to the evaluation set.
+
+These preprocessing tasks were accomplished using Bash and Node.js
+scripts.
+
 \subsection{Classification algorithms}
+
 \subsubsection{Decision tree}

### 2018-01-03 article.tex
Methods

@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage{textcomp}
@@
 \section{Introduction}
-The last decades have witnessed a fast increase on the rate of publication
-of digital text documents. Traditional document types, such as news
-articles, scientific papers or books are now published online together
-with new formats, such as blog posts or tweets, each having thousands
-or millions of new documents published each day.
+Online news articles first appeared as reprints from traditional
+newspapers; nowadays, however, they represent now the primary source
+of news for some segments of the population, both in developed and
+developing countries (whether consumed directly in the newspaper
+website, or indirectly e.g. through a social media application or a
+feed catcher).
+
+Unofficially know as \emph{the fourth branch of government}, the press
+plays a vital role within our society, keeping us informed about
+the current state of affairs (at a local and global scale) and acting
+as a watchdog for the other three branches (legislative, executive and
+judicial). The (lack of) freedom of press and access to the news in a
+given country is even often considered an indicator of a lack of
+democracy\cite{goode2009social}.
+
+As such, improving the ways citizens can access the information (view
+it, query it and search it) contained in news articles has the
+potential to contribute for a more informed and, ultimately, a
+better society.
+
+On the other hand, the last decades have witnessed a fast increase on
+the rate of publication of digital text documents. Traditional
+document types, such as news articles, scientific papers or books are
+now published online together with new formats, such as blog posts or
+tweets, each having thousands or millions of new documents published
+each day.
 
@@
 lifecycle happens digitally, with virtual tools available for
-preparing, writing, styling, publishing and sharing.
+researching, writing, styling, publishing and sharing.
 
@@
 some opportunities when compared to the traditional process. In
-particular, due to todays processing power, tasks related to the
-manipulation of the information contained within these documents
-(searching, compiling, annotating, sharing, \dots) can now be
-performed automatically and targetting a large amount of articles.
+particular, due to the current processing power commonly available,
+tasks related to the manipulation of the information contained within
+these documents (searching, compiling, annotating, sharing, \dots) can
+now be performed automatically and targetting a large amount of
+articles.
 
@@
 
-Online news articles first appeared as reprints from traditional
-newspapers; nowadays, however, they represent now the primary source
-of news for some segments of the population, both in developed and
-developing countries (whether consumed directly in the newspaper
-website, or indirectly (e.g. through a social media application or a
-feed catcher).
-
-Unofficially know as \emph{the fourth branch of government}, the press
-plays a vital role within our society, keeping us informed regarding
-the current state of affairs (at a local and global scale) and acting
-as a watchdog for the other three powers (legislative, executive and
-judicial). The (lack of) freedom of press and access to the news in a
-given country is even often considered an indicator of a lack of
-democracy\cite{goode2009social}.
-
-As such, improving the ways citizens can access the information (view
-it, query it and search it) contained in news articles has the
-potential to contribute for a more informed society and, ultimately, a
-better society.
-
 An example of a feature which improves information access is the
@@
 the information is stored, organized, displayed and queried.
+
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.97\columnwidth]{imgs/wireframe}}
+\caption{Category classification and suggestion on the client side}
+\label{fig:wireframe}
+\end{figure}
 
@@
 Thus, an automated way of categorizing news articles could solve some
-of these problems and decrease the burden of this task.
+of these problems and decrease the burden of this task. Additionally,
+a lightweight version of such a classifier could be implemented on the
+client side code of a web application, for example, allowing the
+categorization to happen in real time (i.e. as the author types in the
+article text). Figure~\ref{fig:wireframe} presents a suggestion of
+how this feature could look like if implemented on a web application.
 
@@
 In this article, we describe the preliminary results obtained in
-developing a classifier to categorize news articles, and in developing
-a similar classifier with restrictions added to the maximum size of
-the lexicon used % TODO
+developing a classifier to categorize news articles using a previously
+manually categorized corpus. Additionally, we evaluate the possibility
+of implementing such a classifier as lightweight as possible to allow
+it to run on the client side of a web application.
 
@@
 \textit{Observador}\footnote{\url{http://observador.pt}}, one of the
-main Portuguese newspapers, and which stands out from the others for
-being fairly young (it was created in May 2014) and for existing
-exclusively online. The initial dataset comprised 42.475 entries, from
-which we used only a subset, for reasons later described.
+main Portuguese newspapers, which stands out from the others for being
+fairly young (it was created in May 2014) and for existing exclusively
+online. The initial dataset comprised 42.475 entries, the most recent
+ones dated from November 2016, from which we used only a subset, for
+reasons later described.
 
@@
 from the most common to the least common. We selected the ones which
-had more than 1.000 articles in our dataset, and reduced our original
-dataset to include only articles from these categories.
+had more than 1000 articles in our dataset, and reduced our original
+dataset to include articles from these categories only.
 Figure~\ref{fig:articles} presents an overview of the selected
@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.97\columnwidth]{imgs/articles}}
-\caption{Total number of articles retrieved for each category}
-\label{fig:articles}
-\end{figure}
-
-We them randomly selected, from each category, 700 articles to be used
-to train the classifiers, and 200 to be used to evaluate their
-performance.
-
-For each article, we had available its contents (title, lead, body)
-and several metadata fields (publication date, category, tags, etc). A
-truncated JSON representation of an article can be found in
-Listing~\ref{lst:article}.
-
-\begin{lstlisting}[caption={Example of JSON representation of an article},label={lst:article}, extendedchars=true]
+\begin{lstlisting}[float,floatplacement=H,caption={Example of JSON representation of an article},label={lst:article}, extendedchars=true]
 {
@@
   URL: "http://observador.pt/2014/08/25/ruben-amorim-com-rotura-total-ligamento-cruzado/",
-  CategoryPaths: [
-      "Desporto"
-  ],
+  CategoryPaths: ["Desporto"],
   Domain: "observador.pt",
@@
 
+\begin{figure}[htbp]
+\centerline{\includegraphics[width=.97\columnwidth]{imgs/articles}}
+\caption{Total number of articles retrieved for each category}
+\label{fig:articles}
+\end{figure}
+
+We them randomly selected, from each category, 700 articles to be used
+to train the classifiers, and 200 to be used to evaluate their
+performance.
+
+For each article, we had available its contents (title, lead, body)
+and several metadata fields (publication date, category, tags, etc). A
+truncated JSON representation of an article can be found in
+Listing~\ref{lst:article}.
+
 \subsection{Preprocessing the articles}
+\label{sec:preproc}
 
-Originally, the dataset was obtained as large (more than 2.5 million
-entries) MongoDB collection, containing articles from several
-Portuguese and international newspapers. The process required to
-transform this collection into data our classifiers could process
-required: querying the database, exporting the news articles and
-splitting them into a train and an evaluation datasets.
+Originally, the dataset was obtained as large MongoDB collection (more
+than 2.5 million entries), containing articles from several Portuguese
+and international newspapers. The process needed to transform this
+collection into data our classifiers could process required querying
+the database, exporting the news articles and splitting them into a
+train and an evaluation datasets.
 
@@
 composed of only a few words), and the categories included at least
-one of the categories we previously selected.
+one of the most common categories.
 
@@
 Furthermore, for each category, 700 articles were allocated to the
-train set, and 200 to the evaluation set.
+train set, and 200 to the evaluation set. These preprocessing tasks
+were accomplished using Bash and Node.js scripts.
 
-These preprocessing tasks were accomplished using Bash and Node.js
-scripts.
+Both datasets were then loaded into R using the
+\texttt{tm}\footnote{\url{https://cran.r-project.org/web/packages/tm/}}
+package. Each was then passed to a function responsible for
+preprocessing the text of the articles:
 
-\subsection{Classification algorithms}
+\begin{itemize}
+    \item the text was converted to lowercase characters
+    \item the Portuguese stopwords were removed
+    \item diacritics were converted to their normalized form (e.g.
+        \framebox{à} $\rightarrow$ \framebox{a})
+    \item punctuation signs were removed
+    \item numbers were removed
+    \item word were stemmed (e.g. \framebox{conseguiram} $\rightarrow$
+        \framebox{consegu})
+    \item whitespace was removed and the text was tokenized
+\end{itemize}
 
-\subsubsection{Decision tree}
-\subsubsection{k-nearest neighbors}
-\subsubsection{Naive Bayes}
-\subsubsection{Neural network}
-\subsubsection{Support vector machine}
+%TODO
+A document-term matrix was then calculated using the
+\texttt{DocumentTermMatrix} method from the \texttt{tm} package. 
+For each document in the train set the matrix included all the
+terms which met the following requirements:
 
-\subsection{Most informative terms}
-% for real-time topic suggestion (for an IDE for example)
-% memory limit
-% 100 terms? 1000 terms?
+\begin{itemize}
+    \item the term length was between 3 and 30 characters (to discard
+        things like URLs or badly tokenized sentences)
+    \item the term appeared in the document at least twice
+    \item the term appeared at least in 10 documents
+\end{itemize}
+The algorithm used to weigh the terms in the matrix was the
+\textit{tf-idf}, and resulted in 3234 terms.
+
+%TODO matrix stats(sparcity etc)
+% inspect(dtm)
+
+The final step was to remove the sparse terms from the matrix.
+This allows to discard terms which might be too specific of the train
+set and which might negatively influence the performance of the
+classifiers by overfitting them to the train set. Additionally, a
+smaller list of terms reduces the execution time of both training and
+applying the classifier, and the memory footprint of the classifier.
+However, while discarding sparse terms we might end up removing
+relevant terms and increase our classifiers error rate.
+
+We used the \texttt{removeSparseTerms} function from the \texttt{tm}
+package, and produced three distinct lists of terms:
+\begin{description}
+    \item[$TS_{99}$] contained the terms with a sparcity level below 99\% (912 terms)
+    \item[$TS_{95}$] contained terms whose sparcity was under 95\% (139 terms)
+    \item[$TS_{90}$] contained terms with 90\% or less sparcity (45 terms)
+\end{description}
+
+\subsection{Classification}
+
+To create the classifiers, we used 5 well knows algorithms: decision
+tree (DT), k-nearest neighbors (KNN), naive Bayes (NB), neural network
+(NN) and support vector
+machine -- with radial kernel (SVM\textsubscript{RK}) and linear
+kernel (SVM\textsubscript{LK}).
+
+We trained each of the classifiers three different times, one for each
+list of terms ($TS_{99}$, $TS_{95}$ and $TS_{90}$). Then we evaluated
+each trained classifier using the test dataset.
+
+The test dataset was preprocessed in a way similar to the train set
+(described in the previous section,~\ref{sec:preproc}),
+but the final list of terms in each test document was restricted to the
+terms considered for the train set.
 
@@
 \begin{table}[htbp]
+    \caption{Error rates obtained for each algorithm and each list of
+    terms}
+\begin{center}
+\begin{tabular}{l|rrrrrr}
+                 & DT            & KNN           & NB   & NN   & SVM\textsubscript{RK} & SVM\textsubscript{LK} \\ \hline
+    $TS_{90}$    & \textbf{0.60} & 0.71          & 0.85 & 0.61 & 0.88                  & 0.88                  \\
+    $TS_{95}$    & 0.70          & \textbf{0.56} & 0.83 & 0.58 & 0.87                  & 0.87                  \\
+    $TS_{99}$    & 0.70          & \textbf{0.35} & 0.87 & 0.76 & 0.87                  & 0.87                  \\
+\end{tabular}
+\label{error}
+\end{center}
+\end{table}
+\begin{table}[htbp]
     \caption{Categories confusion matrix (k-nearest neighbors)}
@@
 \begin{tabular}{l|rrrrrrrr}
-              & ciencia     & cultura     & desporto     & economia     & mundo       & musica       & pais        & politica \\\hline
+& \rotatebox{75}{ciencia} & \rotatebox{75}{cultura} &
+    \rotatebox{75}{desporto} & \rotatebox{75}{economia} &
+    \rotatebox{75}{mundo} & \rotatebox{75}{musica} &
+    \rotatebox{75}{pais} & \rotatebox{75}{politica} \\\hline
 ciencia       & \textbf{64} & 18          & 25           & 29           & 21          & 37           & 4           & 2 \\
@@
 \end{tabular}
-\label{tab1}
+\label{confmat}
 \end{center}
 \end{table}
-% >\\  error.rate.knn
-% [1] 0.5368156
-\subsection{Evaluation}
-\subsection{Limited lexicon}
 

### 2018-01-03 article.tex
Execution times

@@
 \usepackage{float}
+\usepackage{booktabs}
 \usepackage{textcomp}
@@
 \usepackage{hyperref}
+\usepackage{siunitx}
 \usepackage[T1]{fontenc}
@@
 \begin{description}
-    \item[$TS_{99}$] contained the terms with a sparcity level below 99\% (912 terms)
-    \item[$TS_{95}$] contained terms whose sparcity was under 95\% (139 terms)
-    \item[$TS_{90}$] contained terms with 90\% or less sparcity (45 terms)
+    \item[$LT_{90}$] contained terms with 90\% or less sparcity
+    \item[$LT_{95}$] contained terms whose sparcity was under 95\%
+    \item[$LT_{99}$] contained the terms with a sparcity level below 99\%
 \end{description}
@@
 
+For the process of training a
+
 We trained each of the classifiers three different times, one for each
-list of terms ($TS_{99}$, $TS_{95}$ and $TS_{90}$). Then we evaluated
+list of terms ($LT_{99}$, $LT_{95}$ and $LT_{90}$). Then we evaluated
 each trained classifier using the test dataset.
@@
 
+Table~\ref{tab:size} presents the size of each list of terms after
+removing the
+
+\begin{tabular}{S[table-format=3.0]}
+\toprule
+{Heading} \\
+\midrule
+1 \\
+10 \\
+100 \\
+\bottomrule
+\end{tabular}
+
 \begin{table}[htbp]
-    \caption{Error rates obtained for each algorithm and each list of
-    terms}
+    \caption{Lists size}
+\begin{center}
+\begin{tabular}{l|SS}
+    & \parbox{6em}{\centering Sparsity\\threshold (\%)}  &
+    \parbox{5em}{\centering Number\\of terms}   \\\hline
+    $LT_{90}$ &  90                 &  45 \\
+    $LT_{95}$ &  95                 & 139 \\
+    $LT_{99}$ &  99                 & 912 \\
+\end{tabular}
+\label{tab:size}
+\end{center}
+\end{table}
+
+Table~\ref{tab:error} presents the error rates obtained for each list
+of terms using each of the classifiers, with the value of the best
+classifier for each list highlighted in bold.
+
+\begin{table}[htbp]
+    \caption{Error rates}
 \begin{center}
@@
                  & DT            & KNN           & NB   & NN   & SVM\textsubscript{RK} & SVM\textsubscript{LK} \\ \hline
-    $TS_{90}$    & \textbf{0.60} & 0.71          & 0.85 & 0.61 & 0.88                  & 0.88                  \\
-    $TS_{95}$    & 0.70          & \textbf{0.56} & 0.83 & 0.58 & 0.87                  & 0.87                  \\
-    $TS_{99}$    & 0.70          & \textbf{0.35} & 0.87 & 0.76 & 0.87                  & 0.87                  \\
+    $LT_{90}$    & \textbf{0.60} & 0.71          & 0.85 & 0.61 & 0.88                  & 0.88                  \\
+    $LT_{95}$    & 0.70          & \textbf{0.56} & 0.83 & 0.58 & 0.87                  & 0.87                  \\
+    $LT_{99}$    & 0.70          & \textbf{0.35} & 0.87 & 0.76 & 0.87                  & 0.87                  \\
 \end{tabular}
-\label{error}
+\label{tab:error}
 \end{center}
 \end{table}
+
+Table~\ref{tab:exectime} presents the execution time for training the
+algorithm with the lowest error rate for each list of terms.
+
 \begin{table}[htbp]
-    \caption{Categories confusion matrix (k-nearest neighbors)}
+    \caption{Execution times for training}
+\begin{center}
+\begin{tabular}{lcr}
+                 & Algorithm  & Execution time \\\hline
+    $LT_{90}$    & DT         & 12s  \\
+    $LT_{95}$    & KNN        &  3m  \\
+    $LT_{99}$    & KNN        & 57m  \\
+\end{tabular}
+\label{tab:exectime}
+\end{center}
+\end{table}
+
+Table~\ref{tab:confmat} presents the confusion matrix generated using
+the k-nearest neighbors classifier with the $LT_{99}$ list of therms,
+with the number of correct classifications for each category
+highlighted in bold.
+
+\begin{table}[htbp]
+    \caption{Categories confusion matrix ($LT_{99}$ with KNN)}
 \begin{center}
@@
     \rotatebox{75}{pais}     & \rotatebox{75}{politica} \\\hline
-ciencia       & \textbf{64} & 18          & 25           & 29           & 21          & 37           & 4           & 2 \\
-cultura       & 4           & \textbf{37} & 16           & 9            & 19          & 107          & 3           & 5 \\
-desporto      & 1           & 4           & \textbf{130} & 8            & 19          & 35           & 2           & 1 \\
-economia      & 0           & 6           & 9            & \textbf{127} & 12          & 32           & 1           & 13 \\
-mundo         & 9           & 8           & 25           & 31           & \textbf{82} & 32           & 1           & 12 \\
-musica        & 0           & 6           & 1            & 1            & 0           & \textbf{180} & 0           & 1 \\
-pais          & 8           & 15          & 24           & 42           & 17          & 64           & \textbf{14} & 16 \\
-politica      & 1           & 8           & 7            & 36           & 19          & 24           & 3           & \textbf{102} \\
-
+ciencia  & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
+cultura  & 5         & {\bf 102} & 2          & 2         & 3        & 79        & 6        & 1 \\
+desporto & 1         & 3         & {\bf  168} & 7         & 3        & 14        & 4        & 0 \\
+economia & 5         & 3         & 0          & {\bf 146} & 5        & 17        & 14       & 10 \\
+mundo    & 12        & 12        & 7          & 17        & {\bf 90} & 35        & 16       & 11 \\
+musica   & 0         & 2         & 0          & 0         & 2        & {\bf 184} & 1        & 0 \\
+pais     & 9         & 9         & 9          & 19        & 15       & 31        & {\bf 88} & 21 \\
+politica & 1         & 3         & 0          & 30        & 8        & 17        & 14       & {\bf 127} \\
 \end{tabular}
-\label{confmat}
+\label{tab:confmat}
 \end{center}
@@
 
+Looking to the error rate obtained in each of tests performed, we can
+conclude that the best option was to use the largest list of terms
+($LT_{99}$) and the k-nearest neighbors classifier.
+
 \section*{Acknowledgment}
 
-André Santos has a scholarship from Fundação para a
-Computação Científica Nacional.
+André Santos has a PhD scholarship from Fundação para a Ciência e
+Tecnologia.
 

### 2018-01-03 article.tex
Contributions and Future Work

@@
 \texttt{DocumentTermMatrix} method from the \texttt{tm} package.
-For each document in the train set the matrix included all the
-terms which met the following requirements:
+For each of the 5600 documents preset in the train set the matrix
+included all the terms which met the following requirements:
 
@@
 
-The test dataset was preprocessed in a way similar to the train set
-(described in the previous section,~\ref{sec:preproc}),
-but the final list of terms in each test document was restricted to the
-terms considered for the train set.
+The test dataset contained 1600 documents (200 belonging to each
+category) and was preprocessed in a way similar to the train set
+(described in the previous section,~\ref{sec:preproc}).  In each
+iteration, however, the final list of terms in each test document was
+restricted to terms present in the corresponding list of terms
+($LT_{99}$, $LT_{95}$ and $LT_{90}$).
 
 \section{Results}
+A number of measurements and metrics were calculated regarding the
+lists of terms, the classifiers training process and their results in
+the evaluation process.
 
 Table~\ref{tab:size} presents the size of each list of terms after
-removing the
-
-\begin{tabular}{S[table-format=3.0]}
-\toprule
-{Heading} \\
-\midrule
-1 \\
-10 \\
-100 \\
-\bottomrule
-\end{tabular}
+removing the terms whose sparsity was above the corresponding
+threshold.
 
@@
 \label{tab:size}
+\end{center}
+\end{table}
+
+Table~\ref{tab:exectime} presents the execution time for training the
+algorithm with the lowest error rate for each list of terms.
+
+\begin{table}[htbp]
+    \caption{Execution times for training}
+\begin{center}
+\begin{tabular}{l|cr}
+    & Algorithm  & \parbox{3em}{\centering Training\\time} \\\hline
+    $LT_{90}$    & DT         & 12s  \\
+    $LT_{95}$    & KNN        &  3m  \\
+    $LT_{99}$    & KNN        & 57m  \\
+\end{tabular}
+\label{tab:exectime}
 \end{center}
@@
 
-Table~\ref{tab:exectime} presents the execution time for training the
-algorithm with the lowest error rate for each list of terms.
-
-\begin{table}[htbp]
-    \caption{Execution times for training}
-\begin{center}
-\begin{tabular}{lcr}
-                 & Algorithm  & Execution time \\\hline
-    $LT_{90}$    & DT         & 12s  \\
-    $LT_{95}$    & KNN        &  3m  \\
-    $LT_{99}$    & KNN        & 57m  \\
-\end{tabular}
-\label{tab:exectime}
-\end{center}
-\end{table}
-
 Table~\ref{tab:confmat} presents the confusion matrix generated using
@@
 \begin{tabular}{l|rrrrrrrr}
-& \rotatebox{75}{ciencia}    & \rotatebox{75}{cultura}  &
-    \rotatebox{75}{desporto} & \rotatebox{75}{economia} &
-    \rotatebox{75}{mundo}    & \rotatebox{75}{musica}   &
-    \rotatebox{75}{pais}     & \rotatebox{75}{politica} \\\hline
-ciencia  & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
-cultura  & 5         & {\bf 102} & 2          & 2         & 3        & 79        & 6        & 1 \\
-desporto & 1         & 3         & {\bf  168} & 7         & 3        & 14        & 4        & 0 \\
-economia & 5         & 3         & 0          & {\bf 146} & 5        & 17        & 14       & 10 \\
-mundo    & 12        & 12        & 7          & 17        & {\bf 90} & 35        & 16       & 11 \\
-musica   & 0         & 2         & 0          & 0         & 2        & {\bf 184} & 1        & 0 \\
-pais     & 9         & 9         & 9          & 19        & 15       & 31        & {\bf 88} & 21 \\
-politica & 1         & 3         & 0          & 30        & 8        & 17        & 14       & {\bf 127} \\
+& \rotatebox{75}{science}    & \rotatebox{75}{culture}  &
+    \rotatebox{75}{sports} & \rotatebox{75}{economics} &
+    \rotatebox{75}{international}    & \rotatebox{75}{music}   &
+    \rotatebox{75}{national}     & \rotatebox{75}{politics} \\\hline
+science  & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
+culture  & 5         & {\bf 102} & 2          & 2         & 3        & 79        & 6        & 1 \\
+sports   & 1         & 3         & {\bf  168} & 7         & 3        & 14        & 4        & 0 \\
+economics & 5         & 3         & 0          & {\bf 146} & 5        & 17        & 14       & 10 \\
+international    & 12        & 12        & 7          & 17        & {\bf 90} & 35        & 16       & 11 \\
+music   & 0         & 2         & 0          & 0         & 2        & {\bf 184} & 1        & 0 \\
+national     & 9         & 9         & 9          & 19        & 15       & 31        & {\bf 88} & 21 \\
+politics & 1         & 3         & 0          & 30        & 8        & 17        & 14       & {\bf 127} \\
 \end{tabular}
@@
 \section{Discussion}
-% nested categories
-% or not mutex categories
 
-Looking to the error rate obtained in each of tests performed, we can
-conclude that the best option was to use the largest list of terms
-($LT_{99}$) and the k-nearest neighbors classifier.
+The analysis of the results obtained should take into account other
+metrics besides the error rates obtained for each classifier.
+
+The size of the list of terms, for example, gives us an idea of the
+memory footprint of a classifier, a parameter which is of the utmost
+importance if the goal is to implement a classifier as lightweight as
+possible. The training time is also relevant, as shorter training
+times give the possibility of retraining the classifiers more often,
+allowing them to be updated as the corpus of news articles grows.
+
+Looking at the actual results obtained and represented in
+Tables~\ref{tab:size} and \ref{tab:exectime} we can see that $LT_{90}$
+has simultaneously the smallest list of terms (45) and the shortest
+training time (12 seconds using the decision tree algorithm). However,
+$LT_{90}$ also presents the worst error rates, even when looking at
+the algorithm which achieved its best results (0.60 using a decision
+tree classifier).
+
+On the opposite side, $LT_{99}$ presented the lowest error rates (0.35
+using a k-nearest neighbors classifier), but it took almost one hour
+to train and used a list comprising 912 terms.
+
+The results obtained confirm that there is a tradeoff between the size
+of the list of terms and the training time, on the one hand, and the
+classifier error rate, on the other. However, a list of 912 terms
+seems to be an acceptable memory footprint for a client side
+classifier; additionally, the additional training time would not present
+much problems in this scenario, as the classifier would be trained
+before hand and thus would not be visible on the client side.
+
+Given the reasonable values for the training time and the size of the
+list of terms, and looking to the error rate obtained in each of tests
+performed, we can conclude that the best option was to use the largest
+list of terms ($LT_{99}$) and the k-nearest neighbors classifier.
+
+It is worth noting that the categories of an article are not mutually
+exclusive. In fact, an article can be classified as belonging to more
+than one category. This might explain the greater error rates obtained
+in the categories \textit{national} and  \textit{international}: one
+might argue that these categories correspond less to the topic covered
+by the article and are more related to the location of the news
+content.
+
+\section{Contributions and Future work}
+For copyright reasons, the corpus used to train and evaluate the
+classifiers described in this article cannot be shared. All the code
+used to process the documents, to implement the classifiers and
+evaluate them can be found at\\
+\url{http://github.com/andrefs/mapi-msr-categorization}.
+
+The tasks and results previously described already provide useful
+insights into this matter. However, the lowest error rate obtained
+(0.35), despite already acceptable, might still be improved, either by
+trying new algorithms, fine tuning the ones already tested, or by
+increasing the train corpus.
+
+We have established that it is in fact possible
+to develop a news article classifier which is lightweight enough to be
+used (in real time) on the client side of a web application. The
+following step will be the actual implementation, probably in the form
+of a JavaScript library, of such a classifier.
 

### 2018-01-03 article.tex
references

@@
 
-\begin{abstract} This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper
-[title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special
-Characters, Footnotes, or Math in Paper Title or Abstract.
+\begin{abstract} 
+Categorizing news articles according to their contents allows to
+decrease the information entropy in a world where the rate of
+publication of digital text documents is increasing fast. In this
+article we describe ongoing work which aims to evaluate the
+feasability of implementing a classifier which is lightweight
+enough to be used in real time on the client side of a web
+application.  More specifially, we gathered a corpus of Portuguese
+news and used it to train and evaluate several classification
+algorithms. We analyze the results obtained in terms of the
+classifiers error rate, traning time and memory footprint.
 \end{abstract}
@@
 website, or indirectly e.g. through a social media application or a
-feed catcher).
+feed catcher)\cite{greer2006evolution,boczkowski2005digitizing,chyi1999access}.
 
@@
 given country is even often considered an indicator of a lack of
-democracy\cite{goode2009social}.
+democracy\cite{goode2009social,house2009freedom}.
 
@@
 potential to contribute for a more informed and, ultimately, a
-better society.
+better society\cite{bollinger1988tolerant}.
 
@@
 tweets, each having thousands or millions of new documents published
-each day.
+each day\cite{hilbert2011world,allan2006online}.
 
@@
 lifecycle happens digitally, with virtual tools available for
-researching, writing, styling, publishing and sharing.
+researching, writing, styling, publishing and sharing\cite{williams2009personal}.
 
 Having the entire workflow happening within the digital world presents
-some opportunities when compared to the traditional process. In
-particular, due to the current processing power commonly available,
-tasks related to the manipulation of the information contained within
-these documents (searching, compiling, annotating, sharing, \dots) can
-now be performed automatically and targetting a large amount of
-articles.
+some opportunities when compared to the traditional
+process\cite{o1997comparison}. In particular, due to the current
+processing power commonly available, tasks related to the manipulation
+of the information contained within these documents (searching,
+compiling, annotating, sharing, \dots) can now be performed
+automatically and targetting a large amount of articles.
 
@@
 categorization of news articles by the topic (or topics) of its
-content. The presence of such a categorization may influence the way
-the information is stored, organized, displayed and queried.
+content\cite{kim2006extracting}. The presence of such a categorization
+may influence the way the information is stored, organized, displayed
+and queried.
 
@@
 The algorithm used to weigh the terms in the matrix was the
-\textit{tf-idf}, and resulted in 3234 terms.
+\textit{tf-idf}\cite{christopher2008introduction}, and resulted in 3234 terms.
 

### 2018-01-03 article.tex
References

@@
 The challenges of document classification have been well studied
-within the machine learning research field of study. Given a corpus of
-already classified documents, several algorithms might be applied to
-train a classifier capable of determining the category of additional
-articles.
+within the machine learning research field of
+study\cite{borko1963automatic,sebastiani2002machine,rubin2012statistical}.
+Given a corpus of already classified documents, several algorithms
+might be applied to train a classifier capable of determining the
+category of additional articles.
 
@@
 \section{Methods}
-%TODO
-we first needed to choose and obtain a suitable dataset, and then
-cleaning and preparing it to be used to train the classifiers.
+In order to train and evaluate classification algorithms, we first
+needed to choose and obtain a suitable dataset. Preferably, this
+dataset would contain documents were already categorized, which would
+allows to skip the time and effor-consuming step of manually
+categorizing the articles ourselfs. Once this dataset was chosen and
+obtained, we would then clean and prepare it to be used to train the
+classifiers.
 
@@
 
-%TODO
 A document-term matrix was then calculated using the
@@
 The algorithm used to weigh the terms in the matrix was the
-\textit{tf-idf}\cite{christopher2008introduction}, and resulted in 3234 terms.
-
-%TODO matrix stats(sparcity etc)
-% inspect(dtm)
+\textit{tf-idf}\cite{christopher2008introduction}. The obtained matrix
+presented a sparsity of 99\% and contained 3234 distinct terms.
 

### 2018-01-03 article.tex
References and cheating table spaces to fit

@@
 important: author(s), date of publication, source, topic, mentioned
-entities and their relations, etc. Some of this metadata might be
-filled in and stored along with the document (e.g. \emph{author} and
-\emph{date of publication}); other is usually extracted from the
-document content (e.g. mentioned entities).
+entities and their relations, etc\cite{yaginuma2003metadata,yaginuma2003design}. Some of
+this metadata might be filled in and stored along with the document
+(e.g. \emph{author} and \emph{date of publication}); other is usually
+extracted from the document content (e.g. mentioned
+entities)\cite{vadrevu2005automated}.
 
@@
 may influence the way the information is stored, organized, displayed
-and queried.
+and queried\cite{teitler2008newsstand}.
 
@@
 \begin{table}[htbp]
+    \vspace{-.5cm}
     \caption{Lists size}
@@
 \begin{table}[htbp]
+    \vspace{-.2cm}
     \caption{Execution times for training}
@@
 \begin{table}[htbp]
+    \vspace{-.2cm}
     \caption{Error rates}

### 2018-01-04 article.tex
spellcheck

@@
 article we describe ongoing work which aims to evaluate the
-feasability of implementing a classifier which is lightweight
+feasibility of implementing a classifier which is lightweight
 enough to be used in real time on the client side of a web
-application.  More specifially, we gathered a corpus of Portuguese
+application.  More specifically, we gathered a corpus of Portuguese
 news and used it to train and evaluate several classification
 algorithms. We analyze the results obtained in terms of the
-classifiers error rate, traning time and memory footprint.
+classifiers error rate, training time and memory footprint.
 \end{abstract}
@@
 compiling, annotating, sharing, \dots) can now be performed
-automatically and targetting a large amount of articles.
+automatically and targeting a large amount of articles.
 
@@
 dataset would contain documents were already categorized, which would
-allows to skip the time and effor-consuming step of manually
-categorizing the articles ourselfs. Once this dataset was chosen and
+allows to skip the time and effort-consuming step of manually
+categorizing the articles ourselves. Once this dataset was chosen and
 obtained, we would then clean and prepare it to be used to train the
@@
 \begin{description}
-    \item[$LT_{90}$] contained terms with 90\% or less sparcity
-    \item[$LT_{95}$] contained terms whose sparcity was under 95\%
-    \item[$LT_{99}$] contained the terms with a sparcity level below 99\%
+    \item[$LT_{90}$] contained terms with 90\% or less sparsity
+    \item[$LT_{95}$] contained terms whose sparsity was under 95\%
+    \item[$LT_{99}$] contained the terms with a sparsity level below 99\%
 \end{description}

### 2018-01-04 code/classif.R
Code news and paper

@@ -1,0 +1,317 @@
+install.packages("SnowballC")
+install.packages("tm")
+install.packages("FSelector")
+install.packages("caret")
+
+library(tm)
+library(FSelector)
+library(caret)
+library(RWeka)
+library(doMC)
+registerDoMC(cores = 3)
+
+####################################
+# load the corpora and pre process #
+####################################
+
+ciencia.train = VCorpus(DirSource('../news/train/ciencia'), readerControl = list(reader = readPlain, language='pt'))
+ciencia.test = VCorpus(DirSource('../news/eval/ciencia'), readerControl = list(reader = readPlain, language='pt'))
+
+cultura.train = VCorpus(DirSource('../news/train/cultura'), readerControl = list(reader = readPlain, language='pt'))
+cultura.test = VCorpus(DirSource('../news/eval/cultura'), readerControl = list(reader = readPlain, language='pt'))
+
+desporto.train = VCorpus(DirSource('../news/train/desporto'), readerControl = list(reader = readPlain, language='pt'))
+desporto.test = VCorpus(DirSource('../news/eval/desporto'), readerControl = list(reader = readPlain, language='pt'))
+
+economia.train = VCorpus(DirSource('../news/train/economia'), readerControl = list(reader = readPlain, language='pt'))
+economia.test = VCorpus(DirSource('../news/eval/economia'), readerControl = list(reader = readPlain, language='pt'))
+
+mundo.train = VCorpus(DirSource('../news/train/mundo'), readerControl = list(reader = readPlain, language='pt'))
+mundo.test = VCorpus(DirSource('../news/eval/mundo'), readerControl = list(reader = readPlain, language='pt'))
+
+musica.train = VCorpus(DirSource('../news/train/musica'), readerControl = list(reader = readPlain, language='pt'))
[…]

### 2018-01-04 news/scripts/dump2txt.js

@@ -1,0 +1,42 @@
+const mongo = require('promised-mongo');
+const db = mongo('metacache', ['contents']);
+const fs = require('fs-extra');
+const htmlToText = require('html-to-text');
+const Promise = require('bluebird');
+
+async function dumpDocs(filter, folder, fields){
+  console.log('['+(new Date().toISOString())+'] Dumping',filter,'into '+folder);
+  const _filter = {
+    ...filter,
+    'Source.Name': 'Observador',
+    Type: "sapo.obj.creativework.article",
+    Body: {$exists: true},
+    $where: 'this.Body.length > 100'
+  };
+
+  const docs = await db.contents.find(_filter).toArray();
+  return Promise.mapSeries(docs,
+    doc => {
+      const _fields = fields || ['Pretitle','Title','Subtitle', 'Lead'];
+
+      let txt = _fields.map(f => doc[f]);
+      if(doc.Body)    {
+        const text = htmlToText.fromString(doc.Body);
+        txt.push(text);
+      }
+      return fs.writeFile(folder+'/'+doc._id, txt.join('\n'));
+    });
+}
+
+dumpDocs({'CategoryPaths':'Mundo'}, './_/mundo')
+  .then(() => dumpDocs({'CategoryPaths':'País'},     './_/pais'))
+  .then(() => dumpDocs({'CategoryPaths':'Economia'}, './_/economia'))
+  .then(() => dumpDocs({'CategoryPaths':'Política'}, './_/politica'))
+  .then(() => dumpDocs({'CategoryPaths':'Desporto'}, './_/desporto'))
+  .then(() => dumpDocs({'CategoryPaths':'Ciência'},  './_/ciencia'))
+  .then(() => dumpDocs({'CategoryPaths':'Cultura\n\n\tCinema\n\tMúsica\n\tLivros'}, './_/cultura'))
+  .then(() => dumpDocs({'CategoryPaths':'Música'},   './_/musica'))
+  .then(() => {
+    console.log('Finished');
+    process.exit(0);
+  });

### 2018-01-26 paper/article.tex
camera ready

--- 2018-01-04 article.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
+% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.20 (24-JUN-2015)
+
+% !TeX spellcheck = en-US
+% !TeX encoding = utf8
+% !TeX program = pdflatex
+% !BIB program = bibtex
+% -*- coding:utf-8 mod:LaTeX -*-
+
+%"runningheads" enables:
+%  - page number on page 2 onwards
+%  - title/authors on even/odd pages
+%This is good for other readers to enable proper archiving among other papers and pointing to
+%content. Even if the title page states the title, when printed and stored in a folder, when
+%blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
+%it is good to have title printed on the pages, too.
+\documentclass[runningheads,a4paper]{llncs}[2015/06/24]
+
+%cmap has to be loaded before any font package (such as cfr-lm)
+\usepackage{cmap}
+
+\iftrue % use default-font
+  %better font, similar to the default springer font
+  %cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
+  \usepackage[%
+    rm={oldstyle=false,proportional=true},%
+    sf={oldstyle=false,proportional=true},%
+    tt={oldstyle=false,proportional=true,variable=true},%
+    qt=false%
+  ]{cfr-lm}
+\else
+  %
+  %if more space is needed, exchange cfr-lm by two packages from newtx:
+  \usepackage{newtxtext}
+  \usepackage{newtxmath}
+  \usepackage[zerostyle=b,scaled=.9]{newtxtt}
+\fi
+
+%for demonstration purposes only
+\usepackage[math]{blindtext}
+
+%% If you need packages for other papers,
+%% START COPYING HERE
+%%
+%% COPY ALSO cmap from line 19 - it has to be loaded as the first package
+
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc} %support umlauts in the input
+
+\usepackage{graphicx}
+
+%Set English as language and allow to write hyphenated"=words
+%
+%Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
+%english has to go last to set it as default language
+\usepackage[ngerman,english]{babel}
+%Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
+\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}
+
+%Sorts the citations in the brackets
+%It also allows \cite{refa, refb}. Otherwise, the document does not compile.
+%  Error message: "White space in argument"
+\usepackage{cite}
+
+% backticks (`) are rendered as such in verbatim environment. See https://tex.stackexchange.com/a/341057/9075 for details.
+\usepackage{upquote}
+
+%extended enumerate, such as \begin{compactenum}
+\usepackage{paralist}
+
+%put figures inside a text
+%\usepackage{picins}
+%use
+%\piccaptioninside
+%\piccaption{...}
+%\parpic[r]{\includegraphics ...}
+%Text...
+
+%for easy quotations: \enquote{text}
+\usepackage{csquotes}
+
+%enable margin kerning
+\RequirePackage{iftex}
+\ifPDFTeX
+  \RequirePackage[%
+    final,%
+    tracking=smallcaps,%
+    expansion=alltext,%
+    protrusion=alltext-nott]{microtype}%
+\else
+  \RequirePackage[%
+    final,%
+    protrusion=alltext-nott]{microtype}%
+\fi%
+
+%tweak \url{...}
+\usepackage{url}
+%\urlstyle{same}
+%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
+%DO NOT ACTIVATE -> prevents line breaks
+%\makeatletter
+%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
+%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
+%\makeatother
+
+%diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
+%slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
+%\usepackage{diagbox}
+
+%required for pdfcomment later
+\usepackage{xcolor}
+
+%for listings
+\usepackage{listings}
+\lstset{%
+  basicstyle=\ttfamily,%
+  columns=fixed,%
+  basewidth=.5em,%
+  xleftmargin=0.5cm,%
+  literate={á}{{\'a}}1 {ê}{{\^e}}1 {é}{{\'e}}1,
+  captionpos=b}%
+\renewcommand{\lstlistingname}{List.}
+%Fix counter as described at https://tex.stackexchange.com/a/28334/9075
+\usepackage{chngcntr}
+\AtBeginDocument{\counterwithout{lstlisting}{section}}
+
+%compatibility of packages minted and listings with respect to the numbering of "List." caption
+%source: https://tex.stackexchange.com/a/269510/9075
+\AtBeginEnvironment{listing}{\setcounter{listing}{\value{lstlisting}}}
+\AtEndEnvironment{listing}{\stepcounter{lstlisting}}
+
+%enable nice comments
+%this also loads hyperref
+\usepackage{pdfcomment}
+%enable hyperref without colors and without bookmarks
+\hypersetup{hidelinks,
+  colorlinks=true,
+  allcolors=black,
+  pdfstartview=Fit,
+  breaklinks=true}
+%enables correct jumping to figures when referencing
+\usepackage[all]{hypcap}
+
+\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
+\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}
+
+%compatibality with packages todo, easy-todo, todonotes
+\newcommand{\todo}[1]{\commentatside{#1}}
+%compatiblity with package fixmetodonotes
+\newcommand{\TODO}[1]{\commentatside{#1}}
+
+%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link
+\usepackage[capitalise,nameinlink]{cleveref}
+%Nice formats for \cref
+\crefname{section}{Sect.}{Sect.}
+\Crefname{section}{Section}{Sections}
+\crefname{listing}{\lstlistingname}{\lstlistingname}
+\Crefname{listing}{Listing}{Listings}
+
+%\usepackage[newfloat]{minted}
+
+%define IfPackageLoaded
+\usepackage{ltxcmds}
+\makeatletter
+\newcommand{\IfPackageLoaded}[2]{\ltx@ifpackageloaded{#1}{#2}{}}
+\makeatother
+
+\IfPackageLoaded{minted}{
+  % Line numbers not flowing out of the margin
+  \setminted{numbersep=5pt, xleftmargin=12pt}
+
+  %http://www.jevon.org/wiki/Eclipse_Pygments_Style
+  %\usemintedstyle{eclipse}
+  %
+  %\usemintedstyle{autumn}
+  %\usemintedstyle{rrt}
+  %\usemintedstyle{borland}
+  %\usemintedstyle{friendlygrayscale}
+  \usemintedstyle{friendly}
+
+  % We need to load caption to have a bold font on the label
+  % The other parameters mimic the layout of the LNCS class
+  \usepackage[labelfont=bf,font=small,skip=4pt]{caption}
+  \SetupFloatingEnvironment{listing}{name=List.,within=none}
+}{
+}
+%
+%Following definitions are outside of IfPackageLoaded; inside, they are not visible
+%
+%Intermediate solution for hyperlined refs. See https://tex.stackexchange.com/q/132420/9075 for more information.
+\newcommand{\Vlabel}[1]{\label[line]{#1}\hypertarget{#1}{}}
+\newcommand{\lref}[1]{\hyperlink{#1}{\FancyVerbLineautorefname~\ref*{#1}}}
+
+%If minted is not loaded, provide the environments nevertheless
+\ifcsmacro{listing}{}{
+  \newenvironment{listing}[1][htbp!]{\begin{figure}[#1]}{\end{figure}}
+  \newcounter{listing}
+}
+\ifcsmacro{minted}{}{%
+  \newenvironment{minted}[1][]{
+    \bfseries
+    Package minted was not loaded, so there is no XML code shown.
+
+    In case you load minted, please be sure to \begin{enumerate}[a)]
+      \item Have \href{https://www.python.org/downloads/}{python} and \href{http://pygments.org/download/}{pygments} installed
+      \item Excecute pdflatex using \texttt{-shell-escape}
+    \end{enumerate}
+  }{}
+}
+
+\usepackage{xspace}
+%\newcommand{\eg}{e.\,g.\xspace}
+%\newcommand{\ie}{i.\,e.\xspace}
+\newcommand{\eg}{e.\,g.,\ }
+\newcommand{\ie}{i.\,e.,\ }
+
+%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
+\DeclareFontFamily{U}{MnSymbolC}{}
+\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
+\DeclareFontShape{U}{MnSymbolC}{m}{n}{
+  <-6>    MnSymbolC5
+  <6-7>   MnSymbolC6
+  <7-8>   MnSymbolC7
+  <8-9>   MnSymbolC8
+  <9-10>  MnSymbolC9
+  <10-12> MnSymbolC10
+  <12->   MnSymbolC12%
+}{}
+\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}
+
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
+
+%% END COPYING HERE
+
+\usepackage{siunitx}
 \usepackage{graphicx}
 \usepackage{float}
-\usepackage{booktabs}
-\usepackage{textcomp}
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{siunitx}
-\usepackage[T1]{fontenc}
-\usepackage[utf8x]{inputenc}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \begin{document}
 
-\lstset{basicstyle=\footnotesize\ttfamily,breaklines=true,literate={á}{{\'a}}1 {ê}{{\^e}}1 {é}{{\'e}}1 }
+\title{Topic categorization in Portuguese news articles}
+%If Title is too long, use \titlerunning
+%\titlerunning{Short Title}
 
-\title{Topic categorization in Portuguese news articles\\
-\thanks{Fundação para a Ciência e Tecnologia} }
+%Single insitute
+\author{André F. Santos}
+%If there are too many authors, use \authorrunning
+%\authorrunning{First Author et al.}
+\institute{\textit{CRACS \& INESC-Porto LA} \\ \textit{Faculty
+of Sciences, University of Porto}\\ Porto, Portugal \\
+afs@inesctec.pt}
 
-\author{\IEEEauthorblockN{André F. Santos}
-\IEEEauthorblockA{\textit{CRACS \& INESC-Porto LA} \\ \textit{Faculty
-of Sciences, University of Porto}\\ Porto, Portugal \\
-afs@inesctec.pt} }
+%% Multiple insitutes - ALTERNATIVE to the above
+% \author{%
+%     Firstname Lastname\inst{1} \and
+%     Firstname Lastname\inst{2}
+% }
+%
+%If there are too many authors, use \authorrunning
+%  \authorrunning{First Author et al.}
+%
+%  \institute{
+%      Insitute 1\\
+%      \email{...}\and
+%      Insitute 2\\
+%      \email{...}
+%}
 
@@
 
-\begin{IEEEkeywords} topic categorization, machine learning, text
-mining \end{IEEEkeywords}
+\begin{keywords}
+topic categorization, machine learning, text mining
+\end{keywords}
 
-\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}\label{sec:intro}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Online news articles first appeared as reprints from traditional
@@
 
-Unofficially know as \emph{the fourth branch of government}, the press
+Unofficially known as \emph{the fourth branch of government}, the press
 plays a vital role within our society, keeping us informed about
@@
 As such, improving the ways citizens can access the information (view
-it, query it and search it) contained in news articles has the
-potential to contribute for a more informed and, ultimately, a
-better society\cite{bollinger1988tolerant}.
+it, query it and search for it) contained in news articles has the
+potential to contribute for a more informed and, ultimately, better,
+society\cite{bollinger1988tolerant}.
 
@@
 document types, such as news articles, scientific papers or books are
-now published online together with new formats, such as blog posts or
+now published online along with new formats, such as blog posts or
 tweets, each having thousands or millions of new documents published
@@
 
-And it was not only the publication step which has moved to the
-digital world; in fact, most often nowadays the whole document
-lifecycle happens digitally, with virtual tools available for
+Publication is not the only step which has moved to the digital world;
+in fact, most often nowadays the whole document lifecycle happens
+digitally, with virtual tools available for
 researching, writing, styling, publishing and sharing\cite{williams2009personal}.
@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.97\columnwidth]{imgs/wireframe}}
+\begin{figure}[H]
+\centerline{\includegraphics[width=.67\columnwidth]{imgs/wireframe}}
 \caption{Category classification and suggestion on the client side}
@@
 The simplest way of achieving this categorization is to have the
-author of the article to manually introduce it (e.g. the journalist
+author of the article manually introducing it (e.g. the journalist
 typing it on the news article authoring framework); however, this
@@
 \begin{itemize}
-    \item it increases the amount of work the author has to do
-    \item the author might not be sure which categories are available
-    \item the author might not be sure which category is the best
-        (e.g. \emph{Economics} vs  \emph{Finance})
-    \item it does not scale -- e.g. if the goal is to categorize an
-        existing (large) corpus
+    \item It increases the amount of work the author has to do.
+    \item The author might not be sure which categories are available.
+    \item The author might not be sure which category is the best
+        (e.g. \emph{Economics} vs  \emph{Finance}).
+    \item It does not scale -- e.g. if the goal is to categorize an
+        existing (large) corpus.
 \end{itemize}
@@
 needed to choose and obtain a suitable dataset. Preferably, this
-dataset would contain documents were already categorized, which would
-allows to skip the time and effort-consuming step of manually
-categorizing the articles ourselves. Once this dataset was chosen and
-obtained, we would then clean and prepare it to be used to train the
-classifiers.
+dataset should contain documents which were previously categorized,
+allowing us to skip the time and effort-consuming task of
+manually categorizing the articles ourselves. Once this dataset was
+chosen and obtained, we would then clean and prepare it to be used to
+train the classifiers.
 
@@
 categories and the number of articles available for each one.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=.97\columnwidth]{imgs/articles}}
+\caption{Total number of articles retrieved for each category}
+\label{fig:articles}
+\end{figure}
+
+We them randomly selected, from each category, 700 articles to be used
+to train the classifiers, and 200 to be used to evaluate their
+performance.
+
+For each article, we had available its contents (title, lead, body)
+and several metadata fields (publication date, category, tags, etc). A
+truncated JSON representation of an article can be found in
+Listing~\ref{lst:article}.
 
@@
 
-\begin{figure}[htbp]
-\centerline{\includegraphics[width=.97\columnwidth]{imgs/articles}}
-\caption{Total number of articles retrieved for each category}
-\label{fig:articles}
-\end{figure}
-
-We them randomly selected, from each category, 700 articles to be used
-to train the classifiers, and 200 to be used to evaluate their
-performance.
-
-For each article, we had available its contents (title, lead, body)
-and several metadata fields (publication date, category, tags, etc). A
-truncated JSON representation of an article can be found in
-Listing~\ref{lst:article}.
-
 \subsection{Preprocessing the articles}
@@
 
-Originally, the dataset was obtained as large MongoDB collection (more
+Originally, the dataset was obtained as a large MongoDB collection (more
 than 2.5 million entries), containing articles from several Portuguese
@@
 
-The files were then stored in folders, separated by the category.
-Furthermore, for each category, 700 articles were allocated to the
-train set, and 200 to the evaluation set. These preprocessing tasks
-were accomplished using Bash and Node.js scripts.
+The files were stored in folders, separated by category. For each
+category, 700 articles were allocated to the train set, and 200 to the
+evaluation set. These preprocessing tasks were accomplished using Bash
+and Node.js scripts.
 
-Both datasets were then loaded into R using the
+Both datasets were loaded into R using the
 \texttt{tm}\footnote{\url{https://cran.r-project.org/web/packages/tm/}}
@@
 \begin{itemize}
-    \item the text was converted to lowercase characters
-    \item the Portuguese stopwords were removed
-    \item diacritics were converted to their normalized form (e.g.
-        \framebox{à} $\rightarrow$ \framebox{a})
-    \item punctuation signs were removed
-    \item numbers were removed
-    \item word were stemmed (e.g. \framebox{conseguiram} $\rightarrow$
-        \framebox{consegu})
-    \item whitespace was removed and the text was tokenized
+    \item The text was converted to lowercase characters.
+    \item The Portuguese stopwords were removed.
+    \item Diacritics were converted to their normalized form (e.g.
+        \framebox{à} $\rightarrow$ \framebox{a}).
+    \item Punctuation signs were removed.
+    \item Numbers were removed.
+    \item Word were stemmed (e.g. \framebox{conseguiram} $\rightarrow$
+        \framebox{consegu}).
+    \item White space was removed and the text was tokenized.
 \end{itemize}
 
-A document-term matrix was then calculated using the
-\texttt{DocumentTermMatrix} method from the \texttt{tm} package.
-For each of the 5600 documents preset in the train set the matrix
-included all the terms which met the following requirements:
+A document-term matrix was calculated using the
+\texttt{DocumentTermMatrix} method from the \texttt{tm} package.  For
+each of the 5600 documents present in the train set the matrix included
+all the terms which met the following requirements:
 
 \begin{itemize}
-    \item the term length was between 3 and 30 characters (to discard
-        things like URLs or badly tokenized sentences)
-    \item the term appeared in the document at least twice
-    \item the term appeared at least in 10 documents
+    \item The term length was between 3 and 30 characters (to discard
+        things like URLs or badly tokenized sentences).
+    \item The term appeared in the document at least twice.
+    \item The term appeared at least in 10 documents.
 \end{itemize}
-The algorithm used to weigh the terms in the matrix was the
+The algorithm used to weigh the terms in the matrix was
 \textit{tf-idf}\cite{christopher2008introduction}. The obtained matrix
@@
 \begin{description}
-    \item[$LT_{90}$] contained terms with 90\% or less sparsity
-    \item[$LT_{95}$] contained terms whose sparsity was under 95\%
-    \item[$LT_{99}$] contained the terms with a sparsity level below 99\%
+    \item[-- $LT_{90}$] contained terms with 90\% or less sparsity.
+    \item[-- $LT_{95}$] contained terms whose sparsity was under 95\%.
+    \item[-- $LT_{99}$] contained the terms with a sparsity level
+        below 99\%.
 \end{description}
@@
 
-To create the classifiers, we used 5 well knows algorithms: decision
+To create the classifiers, we used 5 well known algorithms: decision
 tree (DT), k-nearest neighbors (KNN), naive Bayes (NB), neural network
@@
 kernel (SVM\textsubscript{LK}).
-
-For the process of training a
 
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
     \vspace{-.5cm}
@@
     & \parbox{6em}{\centering Sparsity\\threshold (\%)}  &
-    \parbox{5em}{\centering Number\\of terms}   \\\hline
+    \parbox{5em}{\centering Number of\\terms left}   \\\hline
     $LT_{90}$ &  90                 &  45 \\
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
     \vspace{-.2cm}
@@
 
-\begin{table}[htbp]
+\begin{table}[H]
     \caption{Categories confusion matrix ($LT_{99}$ with KNN)}
@@
 \begin{tabular}{l|rrrrrrrr}
-& \rotatebox{75}{science}    & \rotatebox{75}{culture}  &
-    \rotatebox{75}{sports} & \rotatebox{75}{economics} &
-    \rotatebox{75}{international}    & \rotatebox{75}{music}   &
-    \rotatebox{75}{national}     & \rotatebox{75}{politics} \\\hline
+& \rotatebox{45}{science}        & \rotatebox{45}{culture}   &
+  \rotatebox{45}{sports}         & \rotatebox{45}{economics} &
+  \rotatebox{45}{international}  & \rotatebox{45}{music}     &
+  \rotatebox{45}{national}       & \rotatebox{45}{politics} \\\hline
 science       & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
@@
 times give the possibility of retraining the classifiers more often,
-allowing them to be updated as the corpus of news articles grows.
+allowing them to be updated as the corpus of articles grows in size.
 
@@
 has simultaneously the smallest list of terms (45) and the shortest
-training time (12 seconds using the decision tree algorithm). However,
-$LT_{90}$ also presents the worst error rates, even when looking at
-the algorithm which achieved its best results (0.60 using a decision
-tree classifier).
+training time (12 seconds using the DT algorithm). However, $LT_{90}$
+also presents the worst error rates, even when looking at the
+algorithm which achieved its best results (0.60 using a DT
+classifier).
 
@@
 seems to be an acceptable memory footprint for a client side
-classifier; additionally, the additional training time would not present
+classifier; additionally, the higher training time would not present
 much problems in this scenario, as the classifier would be trained
-before hand and thus would not be visible on the client side.
+beforehand and thus not be visible on the client side.
 
@@
 performed, we can conclude that the best option was to use the largest
-list of terms ($LT_{99}$) and the k-nearest neighbors classifier.
+list of terms ($LT_{99}$) and the KNN classifier.
 
@@
 insights into this matter. However, the lowest error rate obtained
-(0.35), despite already acceptable, might still be improved, either by
-trying new algorithms, fine tuning the ones already tested, or by
-increasing the train corpus.
+(0.35) might still be improved upon, either by leveraging new
+algorithms, fine tuning the ones already tested, or by increasing the
+train corpus.
 
@@
 used (in real time) on the client side of a web application. The
-following step will be the actual implementation, probably in the form
-of a JavaScript library, of such a classifier.
+following step will be the actual implementation of the classifier,
+probably in the form of a JavaScript library.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{splncs03}
 \bibliography{article}
-\bibliographystyle{splncs03}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
